ms_0_intro

### 2018-03-14 man/ms_0_intro.Rmd
update

Drought is a natural phenomenon occurring when water availability is significantly below-normal levels [@Dai2011]. It is one of the most frequent and widespread climatic extremes affecting ecological systems across every climatic zone worldwide [@Mishra2010; @Dai2011; @Ummenhofer2017]. Summer drought is a characteristic feature of the Mediterranean climate [@Lionello2012], but an increase in the frequency and severity of drought events have been recorded for the Mediterranean region [@Hoerling2011], particularly for southern Europe [@VicenteSerrano2014; @Spinoni2015; @Stagge2017] where a trend towards drier summers has been reported for the last three decades [@Spinoni2017b]. Climate change projections suggest that extreme events, like droughts, will become more frequent and severe [@Trenberth2014; @IPCC2013]  @Spinoni2017a], specially in the Mediterranean area [@Spinoni2017a].

@@

Several studies analyzed the drought effects on Mediterranean tree species representing the southermost limit of their distribution [*e.g*. @Pasho2011; @Camarero2011; @SanchezSalguero2012; @Linares2014; @Dorado2017b; @SanchezSalguero2017; @Caminero2018; ...], however only a few works have focused on the resilience to drought of rear-edge populations [*e.g.* @SanchezSalguero2013; @Herrero2014]. Resilience is the capacity of an ecosystem, community or individual to recover pre-disturbance structure and function after a disturbance (@Holling1973; @Lloret2011; @Hodgson2015). The assessment of forest resilience to climatic disturbances, such as extreme droughts, provides critical information about the capacity of forests to maintain their structure and to continue providing valuable ecosystem services. This is particularly relevant for populations located in their rear-edge of their distribution, especially when they are located in Mediterranean mountains.

In this work we combined remote sensing information and dendroecological methods to evaluate the drought impacts in canopy greenness (as a proxy to primary growth) and radial growth of a Mediterranean tree species (*Quercus pyrenaica*) located in the rear-edge of their distribution. We also assessed the resilience of these forest to severe drought events.

Sabemos que la sequía afecta al crecimiento y a la productividad (poner ejemplos de ambos trabajos) ..

RS 
Remote sensing can be used for studying droughts impacts on ecosystems [*e.g.* @Deshayes2006; @Zhang2013; @AghaKouchak2015; @McDowell2015; @Norman2016]. Drought impacts on vegetation using remote sensing have been analyzed at different scales [@VicenteSerrano2013; @Panisset2018; @Assal2016]. For Mediterranean area there were studies covering all region [@Gouveia2017] and others focused on local scales [*e.g* @Lloret2007; @Gouveia2012, @Zribi2016].

hablar algo de las ventajas de la RS ...

>>>>

Drought is a natural phenomenon occurring when water availability is significantly below-normal levels [@Dai2011]. It is one of the most frequent and widespread climatic extremes affecting ecological systems across every climatic zone worldwide [@Mishra2010; @Dai2011; @Ummenhofer2017]. Summer drought is a characteristic feature of the Mediterranean climate [@Lionello2012], but an increase in the frequency and severity of drought events have been recorded for the Mediterranean region [@Hoerling2011], particularly for southern Europe [@VicenteSerrano2014; @Spinoni2015; @Stagge2017] where a trend towards drier summers has been reported for the last three decades [@Spinoni2017b]. Climate change projections suggest that extreme events, like droughts, will become more frequent and severe [@Trenberth2014; @IPCC2013], specially in the Mediterranean area [@Spinoni2017a].

@@

Several studies analyzed the drought effects on Mediterranean tree species representing the southermost limit of their distribution [@Pasho2011; @Camarero2011; @SanchezSalguero2012; @Linares2014; @Dorado2017b; @SanchezSalguero2017; @Caminero2018], however only a few works have focused on the resilience to drought of rear-edge populations [*e.g.* @SanchezSalguero2013; @Herrero2014]. Resilience is the capacity of an ecosystem, community or individual to recover pre-disturbance structure and function after a disturbance [@Holling1973; @Lloret2011; @Hodgson2015]. The assessment of forest resilience to climatic disturbances, such as extreme droughts, provides critical information about the capacity of forests to maintain their structure and to continue providing valuable ecosystem services. This is particularly relevant for populations located in their rear-edge of their distribution, especially when they are located in Mediterranean mountains.

In this work we combined remote sensing information and dendroecological methods to evaluate the drought impacts in canopy greenness (as a proxy to primary growth) and tree-radial growth of a Mediterranean tree species (*Quercus pyrenaica*) located in the rear-edge of their distribution.

Remote sensing can be used for studying droughts impacts on ecosystems [@Deshayes2006; @Zhang2013; @AghaKouchak2015; @McDowell2015; @Norman2016]. Drought impacts on vegetation using remote sensing have been analyzed at different scales [@VicenteSerrano2013; @Assal2016], and for Mediterranean area there were studies covering all region [@Gouveia2017] and others focused on local scales [@Lloret2007; @VicenteSerrano2007; @Gouveia2012; @Zribi2016]. Vegetation indices obtained from satellite, such us EVI (*Enhanced Vegetation Index*) are used as a proxy for assessing vegetation functioning and aboveground net primary production [@Huete2002], and changes in the patterns of these indices are considered as an idicator of the forest response to external environmental factors, such us droughts [*e.g.* @VicenteSerrano2013; @Vicca2016].

Tree-ring width is a widely used proxy for tree vitality [@Fritts1976; @Dobbertin2005; @Bhuyan2017] and the analysis of annual-tree ring widths can be used to study changes in growth as response to drought at the individiual tree-scale [@Gazol2018].

Several studies han combindado

Several studies have combined the use of satellite information with dendrochronological information to study the effect of droughts on vegetation

We also assessed the resilience of these forest to severe drought events.
